---
title: "I-SPIE_secondary_analysis"
author: "Ric Campbell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
  pdf_document: default
---

## FOR INTERNAL PROJECT USE ONLY, PLEASE DO NOT SHARE BEYOND PROJECT TEAM

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(data.table)
library(ggplot2)
library(scales)
library(kableExtra)
library(lubridate)

createOutputTableProcedureGroups <- function(table) {
  table_vars <- colnames(table)[!(colnames(table) %in% c("30-Day Period", "N"))]
  
  tables_formula <- as.formula(paste0("`30-Day Period` ~ `", paste(table_vars, collapse = "` + `"), "`"))
  
  output_table <- dcast(table, 
                        tables_formula,
                        fill = 0L,
                        drop = FALSE,
                        value.var = "N")[order(`30-Day Period`)]
  
  row_spec(column_spec(kable_styling(kbl(output_table), bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = TRUE), 12, bold = TRUE), 16, bold = TRUE)
}

createOutputTableProcedure <- function(table) {
  table_vars <- colnames(table)[!(colnames(table) %in% c("30-Day Period", "N"))]
  
  tables_formula <- as.formula(paste0("`30-Day Period` ~ `", paste(table_vars, collapse = "` + `"), "`"))
  
  output_table <- dcast(table, 
                        tables_formula,
                        fill = 0L,
                        drop = FALSE,
                        value.var = "N")[order(`30-Day Period`)]
  
  row_spec(column_spec(kable_styling(kbl(output_table), bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = TRUE), 30, bold = TRUE), 16, bold = TRUE)
}

createOutputTableGISub<- function(table) {
  table_vars <- colnames(table)[!(colnames(table) %in% c("30-Day Period", "N"))]
  
  tables_formula <- as.formula(paste0("`30-Day Period` ~ `", paste(table_vars, collapse = "` + `"), "`"))
  
  output_table <- dcast(table, 
                        tables_formula,
                        fill = 0L,
                        drop = FALSE,
                        value.var = "N")[order(`30-Day Period`)]
  
  row_spec(column_spec(kable_styling(kbl(output_table), bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = TRUE), 6, bold = TRUE), 16, bold = TRUE)
}


## Read in IE cohort

ie_cohort <- readRDS("D:/I-SPIEDM/I-SPIEDM/data/datasets/IE_cohort_risk_discharge_causal_180_UoS_2021-11-29.rds")


## Change ADMIDATE to date type
  
  ie_cohort[, ADMIDATE := as.Date(ADMIDATE, format = "%Y-%m-%d")]


####~~~~ Read in different file dependent on selection of procedures want to look at ~~~~~~####    

invasive_procedures <- readRDS("D:/I-SPIEDM/I-SPIEDM/data/datasets/first_invasive_procedure_per_procedure_per_date_patient_heart_rule2022-06-16-132547.rds")


## Reduce to only IE cases of interest, and create admission month field

ie_cohort <- ie_cohort[index_admission == TRUE]
ie_cohort[, admission_month := floor_date(ADMIDATE, unit = "month")]


## Create procedure year field and month field

invasive_procedures[, procedure_year := floor_date(ADMIDATE, unit = "year")]
invasive_procedures[, procedure_month := floor_date(ADMIDATE, unit = "month")]


## Give all IE and invasive procedures an ID (first half of id name will be added below)

ie_cohort[, id := 1:.N]
invasive_procedures[, id := 1:.N]

## Change names so can tell fields apart

setnames(ie_cohort, setdiff(colnames(ie_cohort), "ENCRYPTED_HESID"),
         paste("ie", setdiff(colnames(ie_cohort), "ENCRYPTED_HESID"), sep = "_"))

setnames(invasive_procedures, setdiff(colnames(invasive_procedures), "ENCRYPTED_HESID"),
         paste("invasive", setdiff(colnames(invasive_procedures), "ENCRYPTED_HESID"), sep = "_"))


## Merge together IE cases with all invasive procedures (are we interested in IE admissions with no invasive procedures)

ie_invasive <- merge(ie_cohort,
                     invasive_procedures,
                     by = "ENCRYPTED_HESID",
                     all = FALSE)


## Create field with time between IE admission and invasive procedure  

ie_invasive[, days_to_admission := as.integer(invasive_ADMIDATE - ie_ADMIDATE)]


## Retain only records that had invasive procedure prior to IE admission (do we want ones that are same day?)

invasive_before_ie <- ie_invasive[days_to_admission <= 0]


## Retain only records that have an invasive procedure that follows the heart rule (certain procedures can't be on same admissions as heart valve repair/replace)

invasive_before_ie_heart_rule <- invasive_before_ie[invasive_valid_procedure_including_heart_rule == TRUE]


## Create month splits for time of invasive procedure in relation to IE admission, one for with and one for without day of admissions

invasive_before_ie_heart_rule[, months_to_admission_with_admi_day := ceiling(days_to_admission / 30) - 1L]
invasive_before_ie_heart_rule[days_to_admission != 0
                              , months_to_admission_without_admi_day := ceiling((days_to_admission + 1L) / 30) - 1L]


## No need to reduce IE admissions as have 10 years of HES data prior to first IE admission in 2010

## Remove any records that are more than 15 months prior to IE admission
## Have to create two table as they are different limits and cannot limit in table creation below as need them as characters

invasive_before_ie_heart_rule_with <- invasive_before_ie_heart_rule[months_to_admission_with_admi_day >= -15L]

invasive_before_ie_heart_rule_without <- invasive_before_ie_heart_rule[months_to_admission_without_admi_day >= -15L]

```

```{r create_invasive_procedure_table, echo = FALSE}

  invasive_procedures_date_restricted <- invasive_procedures[invasive_procedure_year >= as.Date("2009-01-01")]

  invasive_procedures_count <- invasive_procedures_date_restricted[, .N]
  invasive_procedures_heart_count <- invasive_procedures_date_restricted[invasive_valid_procedure_including_heart_rule == TRUE, .N]
  
  ## Look at number of invasive procedures groups by year
  
  invasive_procedure_groups_per_year <- invasive_procedures_date_restricted[, .N, by = .(invasive_procedure_year,
                                                                                   invasive_procedure_group)][order(invasive_procedure_year,
                                                                                                                    invasive_procedure_group)]
  
  invasive_procedure_groups_year_output <- dcast(invasive_procedure_groups_per_year,
                                                 invasive_procedure_year ~ invasive_procedure_group,
                                                 fill = 0L,
                                                 drop = FALSE,
                                                 value.var = "N")
  
  
  ## Look at all invasive procedure types by year
  
  invasive_procedures_per_year <- invasive_procedures_date_restricted[, .N, by = .(invasive_procedure_year,
                                                                                         invasive_procedure)][order(invasive_procedure_year,
                                                                                                                    invasive_procedure)]
  
  invasive_procedure_year_output <- dcast(invasive_procedures_per_year,
                                          invasive_procedure_year ~ invasive_procedure,
                                          fill = 0L,
                                          drop = FALSE,
                                          value.var = "N")
  
  
  ## Look at number of invasive procedure types by month
  
  invasive_procedure_per_month <- invasive_procedures_date_restricted[, .N, by = .(invasive_procedure_month,
                                                                                   invasive_procedure)][order(invasive_procedure_month,
                                                                                                              invasive_procedure)]
  
  invasive_procedure_month_output <- dcast(invasive_procedure_per_month,
                                          invasive_procedure_month ~ invasive_procedure,
                                          fill = 0L,
                                          drop = FALSE,
                                          value.var = "N")
  
  
  

```


```{r create_tables_broad, echo = FALSE}

invasive_before_ie_heart_rule[, months_to_admission_with_admi_day := as.character(months_to_admission_with_admi_day)]
invasive_before_ie_heart_rule[, months_to_admission_without_admi_day := as.character(months_to_admission_with_admi_day)]

broad_procedure_group_by_month_admi_inc <- cube(invasive_before_ie_heart_rule_with[ie_condition == "Infective Endocarditis broad",
                                                                                   .N, by = .(month = factor(months_to_admission_with_admi_day, 
                                                                                                             levels = c("-15","-14", "-13", "-12", "-11",
                                                                                                                        "-10", "-9", "-8", "-7", "-6", "-5",
                                                                                                                        "-4", "-3", "-2", "-1", NA)),
                                                                                              invasive_procedure_group = factor(invasive_procedure_group,
                                                                                                                                levels = c("cardiac", "ENT", "GI", "GU",
                                                                                                                                           "haematology", "obstetric_gynae",
                                                                                                                                           "renal","respiratory", "skin", "dental", NA)))],
                                                .(sum(N)), by = c("month", "invasive_procedure_group"))


broad_procedure_by_month_admi_inc <- cube(invasive_before_ie_heart_rule_with[ie_condition == "Infective Endocarditis broad",
                                                                             .N, by = .(month = factor(months_to_admission_with_admi_day, 
                                                                                                       levels = c("-15","-14", "-13", "-12", "-11",
                                                                                                                  "-10", "-9", "-8", "-7", "-6", "-5",
                                                                                                                  "-4", "-3", "-2", "-1", NA)),
                                                                                        invasive_procedure = factor(invasive_procedure,
                                                                                                                    levels = c("abortion_dilatation_curetage",
                                                                                                                               "blood_transfusion",
                                                                                                                               "bone_marrow_puncture",
                                                                                                                               "bronchoscopic",
                                                                                                                               "caesarean_delivery",
                                                                                                                               "colonic_surgery",
                                                                                                                               "coronary_angiography",
                                                                                                                               "coronary_artery_bypass",
                                                                                                                               "cystoscopic",
                                                                                                                               "endodontic_procedures",
                                                                                                                               "endoscopic_lower_gi",
                                                                                                                               "endoscopic_prostate",
                                                                                                                               "endoscopic_upper_gi",
                                                                                                                               "ercp",
                                                                                                                               "extraction_surgical_removal_teeth",
                                                                                                                               "haemodialysis_haemofiltration_haemoperfusion",
                                                                                                                               "implantation_cardiac_pacemakers_defibrillators",
                                                                                                                               "nasal_packing_intubation",
                                                                                                                               "other_endoscopic_oesophageal",
                                                                                                                               "other_oral_surgical_procedure",
                                                                                                                               "percutaneous_coronary_procs_stents",
                                                                                                                               "percutaneous_valve_procedures",
                                                                                                                               "peritoneal_dialysis",
                                                                                                                               "restorative_dental_procedure",
                                                                                                                               "scaling_gingival_procedure",
                                                                                                                               "skin_wound_management",
                                                                                                                               "tonsillectomy_adenoidectomy",
                                                                                                                               "transoesophageal_echocardiography",
                                                                                                                               "vaginal_delivery",
                                                                                                                               NA)))],
                                          .(sum(N)), by = c("month", "invasive_procedure"))


broad_procedure_group_by_month_admi_exc <- cube(invasive_before_ie_heart_rule_without[ie_condition == "Infective Endocarditis broad",
                                                                                      .N, by = .(month = factor(months_to_admission_without_admi_day, 
                                                                                                                levels = c("-15","-14", "-13", "-12", "-11",
                                                                                                                           "-10", "-9", "-8", "-7", "-6", "-5",
                                                                                                                           "-4", "-3", "-2", "-1", NA)),
                                                                                                 invasive_procedure_group = factor(invasive_procedure_group,
                                                                                                                                   levels = c("cardiac", "ENT", "GI", "GU",
                                                                                                                                              "haematology", "obstetric_gynae",
                                                                                                                                              "renal","respiratory", "skin", "dental", NA)))],
                                                .(sum(N)), by = c("month", "invasive_procedure_group"))


broad_procedure_by_month_admi_exc <- cube(invasive_before_ie_heart_rule_without[ie_condition == "Infective Endocarditis broad",
                                                                                .N, by = .(month = factor(months_to_admission_without_admi_day, 
                                                                                                          levels = c("-15","-14", "-13", "-12", "-11",
                                                                                                                     "-10", "-9", "-8", "-7", "-6", "-5",
                                                                                                                     "-4", "-3", "-2", "-1", NA)),
                                                                                           invasive_procedure = factor(invasive_procedure,
                                                                                                                       levels = c("abortion_dilatation_curetage",
                                                                                                                                  "blood_transfusion",
                                                                                                                                  "bone_marrow_puncture",
                                                                                                                                  "bronchoscopic",
                                                                                                                                  "caesarean_delivery",
                                                                                                                                  "colonic_surgery",
                                                                                                                                  "coronary_angiography",
                                                                                                                                  "coronary_artery_bypass",
                                                                                                                                  "cystoscopic",
                                                                                                                                  "endodontic_procedures",
                                                                                                                                  "endoscopic_lower_gi",
                                                                                                                                  "endoscopic_prostate",
                                                                                                                                  "endoscopic_upper_gi",
                                                                                                                                  "ercp",
                                                                                                                                  "extraction_surgical_removal_teeth",
                                                                                                                                  "haemodialysis_haemofiltration_haemoperfusion",
                                                                                                                                  "implantation_cardiac_pacemakers_defibrillators",
                                                                                                                                  "nasal_packing_intubation",
                                                                                                                                  "other_endoscopic_oesophageal",
                                                                                                                                  "other_oral_surgical_procedure",
                                                                                                                                  "percutaneous_coronary_procs_stents",
                                                                                                                                  "percutaneous_valve_procedures",
                                                                                                                                  "peritoneal_dialysis",
                                                                                                                                  "restorative_dental_procedure",
                                                                                                                                  "scaling_gingival_procedure",
                                                                                                                                  "skin_wound_management",
                                                                                                                                  "tonsillectomy_adenoidectomy",
                                                                                                                                  "transoesophageal_echocardiography",
                                                                                                                                  "vaginal_delivery",
                                                                                                                                  NA)))],
                                          .(sum(N)), by = c("month", "invasive_procedure"))

```


```{r upper_lower_GI_intervention_sub_table, echo=FALSE}

    upper_lower_gi_procedures_only <- invasive_before_ie_heart_rule_without[invasive_procedure == "endoscopic_upper_gi" | invasive_procedure == "endoscopic_lower_gi"]

    upper_lower_gi_procedures_only[, invasive_procedure := paste(invasive_procedure, invasive_intervention, sep = "_")]
    
    
    broad_procedure_gi_sub_by_month_admi_exc <- cube(upper_lower_gi_procedures_only[ie_condition == "Infective Endocarditis broad",
                                                                                .N, by = .(month = factor(months_to_admission_without_admi_day, 
                                                                                                          levels = c("-15","-14", "-13", "-12", "-11",
                                                                                                                     "-10", "-9", "-8", "-7", "-6", "-5",
                                                                                                                     "-4", "-3", "-2", "-1", NA)),
                                                                                           invasive_procedure = factor(invasive_procedure,
                                                                                                                       levels = c("endoscopic_upper_gi_FALSE",
                                                                                                                                                   "endoscopic_upper_gi_TRUE",
                                                                                                                                                   "endoscopic_lower_gi_FALSE",
                                                                                                                                                   "endoscopic_lower_gi_TRUE",
                                                                                                                                  NA)))],
                                          .(sum(N)), by = c("month", "invasive_procedure"))
    
    
    
```

```{r create_tables_broad_risk, echo = FALSE}


broad_high_risk_group_by_month_admi_exc <- cube(invasive_before_ie_heart_rule_without[ie_condition == "Infective Endocarditis broad" & ie_risk_at_index == "High",
                                                                                      .N, by = .(month = factor(months_to_admission_without_admi_day, 
                                                                                                                levels = c("-15","-14", "-13", "-12", "-11",
                                                                                                                           "-10", "-9", "-8", "-7", "-6", "-5",
                                                                                                                           "-4", "-3", "-2", "-1", NA)),
                                                                                                 invasive_procedure_group = factor(invasive_procedure_group,
                                                                                                                                   levels = c("cardiac", "ENT", "GI", "GU",
                                                                                                                                              "haematology", "obstetric_gynae",
                                                                                                                                              "renal","respiratory", "skin",
                                                                                                                                              "dental", NA)))],
                                                .(sum(N)), by = c("month", "invasive_procedure_group"))


broad_moderate_risk_group_by_month_admi_exc <- cube(invasive_before_ie_heart_rule_without[ie_condition == "Infective Endocarditis broad" & ie_risk_at_index == "Moderate",
                                                                                          .N, by = .(month = factor(months_to_admission_without_admi_day, 
                                                                                                                    levels = c("-15","-14", "-13", "-12", "-11",
                                                                                                                               "-10", "-9", "-8", "-7", "-6", "-5",
                                                                                                                               "-4", "-3", "-2", "-1", NA)),
                                                                                                     invasive_procedure_group = factor(invasive_procedure_group,
                                                                                                                                       levels = c("cardiac", "ENT", "GI", "GU",
                                                                                                                                                  "haematology", "obstetric_gynae",
                                                                                                                                                  "renal","respiratory", "skin",
                                                                                                                                                  "dental", NA)))],
                                                    .(sum(N)), by = c("month", "invasive_procedure_group"))


broad_low_risk_group_by_month_admi_exc <- cube(invasive_before_ie_heart_rule_without[ie_condition == "Infective Endocarditis broad" & ie_risk_at_index == "Low",
                                                                                     .N, by = .(month = factor(months_to_admission_without_admi_day, 
                                                                                                               levels = c("-15","-14", "-13", "-12", "-11",
                                                                                                                          "-10", "-9", "-8", "-7", "-6", "-5",
                                                                                                                          "-4", "-3", "-2", "-1", NA)),
                                                                                                invasive_procedure_group = factor(invasive_procedure_group,
                                                                                                                                  levels = c("cardiac", "ENT", "GI", "GU",
                                                                                                                                             "haematology", "obstetric_gynae",
                                                                                                                                             "renal","respiratory", "skin",
                                                                                                                                             "dental", NA)))],
                                               .(sum(N)), by = c("month", "invasive_procedure_group"))



broad_high_risk_procedure_by_month_admi_exc <- cube(invasive_before_ie_heart_rule_without[ie_condition == "Infective Endocarditis broad" & ie_risk_at_index == "High",
                                                                                          .N, by = .(month = factor(months_to_admission_without_admi_day, 
                                                                                                                    levels = c("-15","-14", "-13", "-12", "-11",
                                                                                                                               "-10", "-9", "-8", "-7", "-6", "-5",
                                                                                                                               "-4", "-3", "-2", "-1", NA)),
                                                                                                     invasive_procedure = factor(invasive_procedure,
                                                                                                                                 levels = c("abortion_dilatation_curetage",
                                                                                                                                            "blood_transfusion",
                                                                                                                                            "bone_marrow_puncture",
                                                                                                                                            "bronchoscopic",
                                                                                                                                            "caesarean_delivery",
                                                                                                                                            "colonic_surgery",
                                                                                                                                            "coronary_angiography",
                                                                                                                                            "coronary_artery_bypass",
                                                                                                                                            "cystoscopic",
                                                                                                                                            "endodontic_procedures",
                                                                                                                                            "endoscopic_lower_gi",
                                                                                                                                            "endoscopic_prostate",
                                                                                                                                            "endoscopic_upper_gi",
                                                                                                                                            "ercp",
                                                                                                                                            "extraction_surgical_removal_teeth",
                                                                                                                                        "haemodialysis_haemofiltration_haemoperfusion",
                                                                                                                                      "implantation_cardiac_pacemakers_defibrillators",
                                                                                                                                            "nasal_packing_intubation",
                                                                                                                                            "other_endoscopic_oesophageal",
                                                                                                                                            "other_oral_surgical_procedure",
                                                                                                                                            "percutaneous_coronary_procs_stents",
                                                                                                                                            "percutaneous_valve_procedures",
                                                                                                                                            "peritoneal_dialysis",
                                                                                                                                            "restorative_dental_procedure",
                                                                                                                                            "scaling_gingival_procedure",
                                                                                                                                            "skin_wound_management",
                                                                                                                                            "tonsillectomy_adenoidectomy",
                                                                                                                                            "transoesophageal_echocardiography",
                                                                                                                                            "vaginal_delivery",
                                                                                                                                            NA)))],
                                                    .(sum(N)), by = c("month", "invasive_procedure"))



broad_moderate_risk_procedure_by_month_admi_exc <- cube(invasive_before_ie_heart_rule_without[ie_condition == "Infective Endocarditis broad" & ie_risk_at_index == "Moderate",
                                                                                              .N, by = .(month = factor(months_to_admission_without_admi_day, 
                                                                                                                        levels = c("-15","-14", "-13", "-12", "-11",
                                                                                                                                   "-10", "-9", "-8", "-7", "-6", "-5",
                                                                                                                                   "-4", "-3", "-2", "-1", NA)),
                                                                                                         invasive_procedure = factor(invasive_procedure,
                                                                                                                                     levels = c("abortion_dilatation_curetage",
                                                                                                                                                "blood_transfusion",
                                                                                                                                                "bone_marrow_puncture",
                                                                                                                                                "bronchoscopic",
                                                                                                                                                "caesarean_delivery",
                                                                                                                                                "colonic_surgery",
                                                                                                                                                "coronary_angiography",
                                                                                                                                                "coronary_artery_bypass",
                                                                                                                                                "cystoscopic",
                                                                                                                                                "endodontic_procedures",
                                                                                                                                                "endoscopic_lower_gi",
                                                                                                                                                "endoscopic_prostate",
                                                                                                                                                "endoscopic_upper_gi",
                                                                                                                                                "ercp",
                                                                                                                                                "extraction_surgical_removal_teeth",
                                                                                                                                        "haemodialysis_haemofiltration_haemoperfusion",
                                                                                                                                      "implantation_cardiac_pacemakers_defibrillators",
                                                                                                                                                "nasal_packing_intubation",
                                                                                                                                                "other_endoscopic_oesophageal",
                                                                                                                                                "other_oral_surgical_procedure",
                                                                                                                                                "percutaneous_coronary_procs_stents",
                                                                                                                                                "percutaneous_valve_procedures",
                                                                                                                                                "peritoneal_dialysis",
                                                                                                                                                "restorative_dental_procedure",
                                                                                                                                                "scaling_gingival_procedure",
                                                                                                                                                "skin_wound_management",
                                                                                                                                                "tonsillectomy_adenoidectomy",
                                                                                                                                                "transoesophageal_echocardiography",
                                                                                                                                                "vaginal_delivery",
                                                                                                                                                NA)))],
                                                        .(sum(N)), by = c("month", "invasive_procedure"))


broad_low_risk_procedure_by_month_admi_exc <- cube(invasive_before_ie_heart_rule_without[ie_condition == "Infective Endocarditis broad" & ie_risk_at_index == "Low",
                                                                                         .N, by = .(month = factor(months_to_admission_without_admi_day, 
                                                                                                                   levels = c("-15","-14", "-13", "-12", "-11",
                                                                                                                              "-10", "-9", "-8", "-7", "-6", "-5",
                                                                                                                              "-4", "-3", "-2", "-1", NA)),
                                                                                                    invasive_procedure = factor(invasive_procedure,
                                                                                                                                levels = c("abortion_dilatation_curetage",
                                                                                                                                           "blood_transfusion",
                                                                                                                                           "bone_marrow_puncture",
                                                                                                                                           "bronchoscopic",
                                                                                                                                           "caesarean_delivery",
                                                                                                                                           "colonic_surgery",
                                                                                                                                           "coronary_angiography",
                                                                                                                                           "coronary_artery_bypass",
                                                                                                                                           "cystoscopic",
                                                                                                                                           "endodontic_procedures",
                                                                                                                                           "endoscopic_lower_gi",
                                                                                                                                           "endoscopic_prostate",
                                                                                                                                           "endoscopic_upper_gi",
                                                                                                                                           "ercp",
                                                                                                                                           "extraction_surgical_removal_teeth",
                                                                                                                                        "haemodialysis_haemofiltration_haemoperfusion",
                                                                                                                                      "implantation_cardiac_pacemakers_defibrillators",
                                                                                                                                           "nasal_packing_intubation",
                                                                                                                                           "other_endoscopic_oesophageal",
                                                                                                                                           "other_oral_surgical_procedure",
                                                                                                                                           "percutaneous_coronary_procs_stents",
                                                                                                                                           "percutaneous_valve_procedures",
                                                                                                                                           "peritoneal_dialysis",
                                                                                                                                           "restorative_dental_procedure",
                                                                                                                                           "scaling_gingival_procedure",
                                                                                                                                           "skin_wound_management",
                                                                                                                                           "tonsillectomy_adenoidectomy",
                                                                                                                                           "transoesophageal_echocardiography",
                                                                                                                                           "vaginal_delivery",
                                                                                                                                           NA)))],
                                                   .(sum(N)), by = c("month", "invasive_procedure"))




```


```{r create_tables_narrow, echo = FALSE}
narrow_procedure_group_by_month_admi_inc <- cube(invasive_before_ie_heart_rule_with[ie_condition == "Infective Endocarditis narrow",
                                                                                    .N, by = .(month = factor(months_to_admission_with_admi_day, 
                                                                                                              levels = c("-15","-14", "-13", "-12", "-11",
                                                                                                                         "-10", "-9", "-8", "-7", "-6", "-5",
                                                                                                                         "-4", "-3", "-2", "-1", NA)),
                                                                                               invasive_procedure_group = factor(invasive_procedure_group,
                                                                                                                                 levels = c("cardiac", "ENT", "GI", "GU",
                                                                                                                                            "haematology", "obstetric_gynae",
                                                                                                                                            "renal","respiratory", "skin",
                                                                                                                                            "dental", NA)))],
                                                 .(sum(N)), by = c("month", "invasive_procedure_group"))


narrow_procedure_group_by_month_admi_exc <- cube(invasive_before_ie_heart_rule_without[ie_condition == "Infective Endocarditis narrow",
                                                                                       .N, by = .(month = factor(months_to_admission_without_admi_day, 
                                                                                                                 levels = c("-15","-14", "-13", "-12", "-11",
                                                                                                                            "-10", "-9", "-8", "-7", "-6", "-5",
                                                                                                                            "-4", "-3", "-2", "-1", NA)),
                                                                                                  invasive_procedure_group = factor(invasive_procedure_group,
                                                                                                                                    levels = c("cardiac", "ENT", "GI", "GU",
                                                                                                                                               "haematology", "obstetric_gynae",
                                                                                                                                               "renal","respiratory", "skin",
                                                                                                                                               "dental", NA)))],
                                                 .(sum(N)), by = c("month", "invasive_procedure_group"))

```



```{r change_field_names, echo = FALSE}
## First set broad - groups and procedures  
broad_procedure_group_by_month_admi_inc[is.na(as.character(month)), month := "Total"]
broad_procedure_group_by_month_admi_inc[is.na(invasive_procedure_group), invasive_procedure_group := "Total"]

broad_procedure_by_month_admi_inc[is.na(as.character(month)), month := "Total"]
broad_procedure_by_month_admi_inc[is.na(invasive_procedure), invasive_procedure := "Total"]

broad_procedure_group_by_month_admi_exc[is.na(as.character(month)), month := "Total"]
broad_procedure_group_by_month_admi_exc[is.na(invasive_procedure_group), invasive_procedure_group := "Total"]

broad_procedure_by_month_admi_exc[is.na(as.character(month)), month := "Total"]
broad_procedure_by_month_admi_exc[is.na(invasive_procedure), invasive_procedure := "Total"]

broad_procedure_gi_sub_by_month_admi_exc[is.na(as.character(month)), month := "Total"]
broad_procedure_gi_sub_by_month_admi_exc[is.na(invasive_procedure), invasive_procedure := "Total"]



## Second set broad - split by risk, all excluding admi data

broad_high_risk_group_by_month_admi_exc[is.na(as.character(month)), month := "Total"]
broad_high_risk_group_by_month_admi_exc[is.na(invasive_procedure_group), invasive_procedure_group := "Total"]

broad_moderate_risk_group_by_month_admi_exc[is.na(as.character(month)), month := "Total"]
broad_moderate_risk_group_by_month_admi_exc[is.na(invasive_procedure_group), invasive_procedure_group := "Total"]

broad_low_risk_group_by_month_admi_exc[is.na(as.character(month)), month := "Total"]
broad_low_risk_group_by_month_admi_exc[is.na(invasive_procedure_group), invasive_procedure_group := "Total"]

broad_high_risk_procedure_by_month_admi_exc[is.na(as.character(month)), month := "Total"]
broad_high_risk_procedure_by_month_admi_exc[is.na(invasive_procedure), invasive_procedure := "Total"]

broad_moderate_risk_procedure_by_month_admi_exc[is.na(as.character(month)), month := "Total"]
broad_moderate_risk_procedure_by_month_admi_exc[is.na(invasive_procedure), invasive_procedure := "Total"]

broad_low_risk_procedure_by_month_admi_exc[is.na(as.character(month)), month := "Total"]
broad_low_risk_procedure_by_month_admi_exc[is.na(invasive_procedure), invasive_procedure := "Total"]




## First set narrow - groups

narrow_procedure_group_by_month_admi_inc[is.na(as.character(month)), month := "Total"]
narrow_procedure_group_by_month_admi_inc[is.na(invasive_procedure_group), invasive_procedure_group := "Total"]

narrow_procedure_group_by_month_admi_exc[is.na(as.character(month)), month := "Total"]
narrow_procedure_group_by_month_admi_exc[is.na(invasive_procedure_group), invasive_procedure_group := "Total"]



setnames(broad_procedure_group_by_month_admi_inc, c("month", "V1"), c("30-Day Period", "N"))
setnames(broad_procedure_by_month_admi_inc, c("month", "V1"), c("30-Day Period", "N"))
setnames(broad_procedure_group_by_month_admi_exc, c("month", "V1"), c("30-Day Period", "N"))
setnames(broad_procedure_by_month_admi_exc, c("month", "V1"), c("30-Day Period", "N"))
setnames(broad_procedure_gi_sub_by_month_admi_exc, c("month", "V1"), c("30-Day Period", "N"))
setnames(broad_high_risk_group_by_month_admi_exc, c("month", "V1"), c("30-Day Period", "N"))
setnames(broad_moderate_risk_group_by_month_admi_exc, c("month", "V1"), c("30-Day Period", "N"))
setnames(broad_low_risk_group_by_month_admi_exc, c("month", "V1"), c("30-Day Period", "N"))
setnames(broad_high_risk_procedure_by_month_admi_exc, c("month", "V1"), c("30-Day Period", "N"))
setnames(broad_moderate_risk_procedure_by_month_admi_exc, c("month", "V1"), c("30-Day Period", "N"))
setnames(broad_low_risk_procedure_by_month_admi_exc, c("month", "V1"), c("30-Day Period", "N"))

setnames(narrow_procedure_group_by_month_admi_inc, c("month", "V1"), c("30-Day Period", "N"))
setnames(narrow_procedure_group_by_month_admi_exc, c("month", "V1"), c("30-Day Period", "N"))

```

  
## Invasive procedures
The following tables summarises the total number of invasive procedures in the years of interest.
  
Min invasive procedure date: `r paste0(invasive_procedures_date_restricted[, min(invasive_ADMIDATE)], collapse = " ")`  
Max invasive procedure date: `r paste0(invasive_procedures_date_restricted[, max(invasive_ADMIDATE)], collapse = " ")`  
  
Total number of invasive procedure: `r paste0(invasive_procedures_date_restricted[, .N], collapse = " ")`  
Number of invasive procedures removed due to having the same admission date as a heart valve repair or replace:`r paste0((invasive_procedures_date_restricted[,.N] - invasive_procedures_date_restricted[invasive_valid_procedure_including_heart_rule == TRUE,.N]), collapse = " ")`  

```{r invasive_procedure_table, echo=FALSE}
  
kable_styling(kable(invasive_procedure_groups_year_output), bootstrap_options = c("striped", "hover", " condensed"))  

kable_styling(kable(invasive_procedure_year_output), bootstrap_options = c("striped", "hover", " condensed"))

kable_styling(kable(invasive_procedure_month_output), bootstrap_options = c("striped", "hover", " condensed"), fixed_thead = TRUE)


```
  
# Data in detail
**Note:** All tables and graphs have now been created using the first invasive surgical procedure (highest positioned relevant OPCS code) per procedure, per person, per admission date. This means that at most a single admission date can have 29 different procedures attached to it.
  
  

## Broad definition
### Tables - split by both procedure groups and procedures
The following tables summarise the total number of invasive procedures prior to admission meeting the *broad* definition of IE.
The only difference between the two tables is that the first includes any invasive procedures that occurred on the day of IE admission, while the second table does not include invasive procedures that occurred on the same day as IE admission. (All groups consist of 30-day periods).

### With admission date
```{r broad_summary_tables_with, echo=FALSE}

createOutputTableProcedureGroups(broad_procedure_group_by_month_admi_inc)

createOutputTableProcedure(broad_procedure_by_month_admi_inc)

```


### Without admissions date
```{r broad_summary_tables_without, echo=FALSE}
createOutputTableProcedureGroups(broad_procedure_group_by_month_admi_exc)

createOutputTableProcedure(broad_procedure_by_month_admi_exc)
```  


### Additional sub-analysis of upper and lower GI group
This table takes the procedures under upper and lower GI, and splits them by if the procedure has been deemed an intervention or a 'look and see' procedure. Does not include procedures done on admission date
```{r gi_sub_analysis_table_without}

createOutputTableGISub(broad_procedure_gi_sub_by_month_admi_exc)

```

## Broad definition
### Tables - split by risk
The following tables summarise the total number of invasive procedures prior to admission meeting the *broad* definition of IE split by the risk category at the time of admissions.


### High risk - by group
```{r broad_risk_tables, echo=FALSE} 
createOutputTableProcedureGroups(broad_high_risk_group_by_month_admi_exc)
```


### Moderate risk - by group
``` {r}
createOutputTableProcedureGroups(broad_moderate_risk_group_by_month_admi_exc)
```


### Low/Unknown risk - by group
``` {r}
createOutputTableProcedureGroups(broad_low_risk_group_by_month_admi_exc)
```


### High risk - by procedure
``` {r}
createOutputTableProcedure(broad_high_risk_procedure_by_month_admi_exc)
```


### Moderate risk - by procedure
``` {r}
createOutputTableProcedure(broad_moderate_risk_procedure_by_month_admi_exc)
```


### Low/Unknown risk - by procedure
``` {r}
createOutputTableProcedure(broad_low_risk_procedure_by_month_admi_exc)
```



### Graphs
The following graphs summarise the total number of invasive procedures prior to admission meeting the *broad* definition of IE.
The only difference between the two tables is that the first includes any invasive procedures that occurred on the day of IE admission, while the second table does not include invasive procedures that occurred on the same day as IE admission.

```{r broad_graphs, echo = FALSE}

ggplot(broad_procedure_group_by_month_admi_inc[`30-Day Period` != "Total",
                                               .(`30-Day Period` = as.numeric(levels(`30-Day Period`)[`30-Day Period`]),
                                                 invasive_procedure_group, N)],
       aes(x = `30-Day Period`, y = N, colour = invasive_procedure_group)) +
  geom_point() + 
  geom_line() +
  scale_y_continuous(limits = c(0,NA)) +
  scale_x_continuous(breaks = -15:-1,
                     minor_breaks = NULL,
                     name = "30-Day Period prior to IE admission") +
  ggtitle("Number of invasive surgical procedures split by procedure group in 30-day
            periods before admission for IE using the broad definition")


ggplot(broad_procedure_group_by_month_admi_exc[`30-Day Period` != "Total",
                                               .(`30-Day Period` = as.numeric(levels(`30-Day Period`)[`30-Day Period`]),
                                                 invasive_procedure_group, N)],
       aes(x = `30-Day Period`, y = N, colour = invasive_procedure_group)) +
  geom_point() + 
  geom_line() +
  scale_y_continuous(limits = c(0,NA)) +
  scale_x_continuous(breaks = -15:-1,
                     minor_breaks = NULL,
                     name = "30-Day Period prior to IE admission") +
  ggtitle("Number of invasive surgical procedures split by procedure group in 30-day
            periods (excluding day of admission) before admission for IE using the broad definition")

```


## Narrow definition
### Tables
The following tables summarise the total number of invasive procedures prior to admission meeting the *narrow* definition of IE.
The only difference between the two tables is that the first includes any invasive procedures that occurred on the day of IE admission, while the second table does not include invasive procedures that occurred on the same day as IE admission. 

```{r narrow_summary_table, echo = FALSE}

createOutputTableProcedureGroups(narrow_procedure_group_by_month_admi_inc)
createOutputTableProcedureGroups(narrow_procedure_group_by_month_admi_exc)

```

### Graphs
The following graphs summarise the total number of invasive procedures prior to admission meeting the *narrow* definition of IE.
The only difference between the two tables is that the first includes any invasive procedures that occurred on the day of IE admission, while the second table does not include invasive procedures that occurred on the same day as IE admission. 


```{r narrow_graphs, echo = FALSE}

ggplot(narrow_procedure_group_by_month_admi_inc[`30-Day Period` != "Total",
                                                .(`30-Day Period` = as.numeric(levels(`30-Day Period`)[`30-Day Period`]),
                                                  invasive_procedure_group, N)],
       aes(x = `30-Day Period`, y = N, colour = invasive_procedure_group)) +
  geom_point() + 
  geom_line() +
  scale_y_continuous(limits = c(0,NA)) +
  scale_x_continuous(breaks = -15:-1,
                     minor_breaks = NULL,
                     name = "30-Day Period prior to IE admission") +
  ggtitle("Number of invasive surgical procedures split by procedure group in 30-day
            periods before admission for IE using the narrow definition")


ggplot(narrow_procedure_group_by_month_admi_exc[`30-Day Period` != "Total",
                                                .(`30-Day Period` = as.numeric(levels(`30-Day Period`)[`30-Day Period`]),
                                                  invasive_procedure_group, N)],
       aes(x = `30-Day Period`, y = N, colour = invasive_procedure_group)) +
  geom_point() + 
  geom_line() +
  scale_y_continuous(limits = c(0,NA)) +
  scale_x_continuous(breaks = -15:-1,
                     minor_breaks = NULL,
                     name = "30-Day Period prior to IE admission") +
  ggtitle("Number of invasive surgical procedures split by procedure group in 30-day
            periods (excluding day of admission) before admission for IE using the narrow definition")

```

